---
title: ""
author: "Michael Hua"
output:
  pdf_document: default
  html_document: default
---

Filter the dataset for two countries of your choice and extract data from a specific data range (e.g. 2022-01-01 to 2022-12-31 Columns to focus on: location, date, new cases, new deaths (you can choose different columns if you would like to) Calculate the total number of cases and total number of deaths during that time.

This program selects two countries and extracts data between New Zealand and Bulgaria from January 2022 to January 2023.

Search by location, date, new cases, new deaths

```{r}
library(dplyr)
covid_data <- read.csv("owid-covid-data.csv") 

#filters data between 2022-01-01 and 2022-12-31 for the countries Bulgaria and New Zealand
filtered_covid_data <- covid_data %>%
  filter(location %in% c("New Zealand", "Bulgaria") & date >= "2022-01-01" & date <= "2023-01-01") %>% #filters by date and country
  select(location, date, new_cases, new_deaths, total_deaths_per_million) #makes set only include these columns

#finds total cases and deaths accumulated during 2022-01-01 and 2022-12-31
summary_covid_data <- filtered_covid_data %>% 
  group_by(location) %>% #separates the two countries
  summarise(total_cases_filtered = sum(new_cases), total_deaths_filtered = sum(new_deaths)) #sums all new cases and deaths during the period

print(filtered_covid_data)
print(summary_covid_data)
```

1.  **Regional Vaccination Rate - Compare total vaccinations per hundred across continents.**

    1.  **Filter the data set for a specific date.**

    2.  **Group_by() , then get mean vaccination rate.**

    3.  **Plot continents vs. vaccination rates.**

    4.  **Draw some conclusions. Which continents have a higher rate? Why?**

``` {r}
library(dplyr)
library(ggplot2)
covid_data <- read.csv("owid-covid-data.csv")
filtered_data_set = covid_data %>%
  filter(date == "2023-01-01", continent !="") %>%  
  #filters by the specific date, removes data not considered in continents
  group_by(continent) %>%
  summarise(mean_vaccination_rate = mean(total_vaccinations_per_hundred, na.rm = TRUE))
#creates new data frame for grouped variable continent, 
#calculates mean vaccination rate, excluding NA values
print(filtered_data_set)

ggplot(filtered_data_set, aes(x = continent, y = mean_vaccination_rate, fill = continent)) + 
  #adds x and y labels, color-codes continents
  geom_bar(stat = "identity") + #adds the bar in the box plot
  labs(
    title = "Mean Total Vaccinations per Hundred by Continent", 
    x = "Continent", 
    y = "Mean Vaccination Rate per Hundred",
    caption = "Source: covid data from somewhere")

```
Conclusions:
South America has the highest mean vaccination rate per hundred, 
followed by Oceania, North America, Europe, Asia, and Africa.
Some factors of this may include economic differences (some countries must purchase vaccines),
the time period (it's 2023, and vaccination rates differ by the day), and
availability (how easy it is to get the vaccine)
https://english.elpais.com/usa/2022-01-13/how-south-america-became-a-global-leader-in-covid-19-vaccination.html?outputType=amp 